---
title: "PCP_alphaWorkflow"
author: "VanMcG"
date: "January 26, 2019"
output: html_document
---

## 00: R Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sp)
library(ggplot2)
library(ggforce) # to plot circles
library(dplyr)
#library(devtools)
#library(SpatialEpi)
library(gmt) # geodist is in geosphere
library(tidyverse)
#library(sf)
library(geosphere)
library(useful)
library(MonoPoly)
library(akima)
library(polynom)
library(extrafont)
library(reshape2)

# USE TO DOWNLOAD XKCD FONT
download.file("http://simonsoftware.se/other/xkcd.ttf", dest="xkcd.ttf", mode="wb")
system("mkdir ~/.fonts")
system("cp xkcd.ttf  ~/.fonts")
font_import(pattern = "xkcd", prompt=FALSE)
# fonts()
# fonttable()

```

## 01: Input CSV into R
* Read and give headers and datatypes for each (eventually do more sophisticated validation/type detection)

```{r}
remits <- read.csv(file = "testData/IND_remittances.csv")

remits <- remits[!(remits$lat>50 | remits$long < -125),]
summary(remits)
class(remits) # test to make sure it's a data frame

# at some point, we need to enforce generic column headers

typeof(remits$isCTR)

     # obnoxiously explicit validation

if (typeof(remits$name) != "character")
  remits$name <- as.character(remits$name)
if (typeof(remits$isCTR) != "logical")
  remits$isCTR <- as.logical(remits$isCTR)
if (typeof(remits$lat) != "double")
  remits$isCTR <- as.double(remits$lat)
if (typeof(remits$long) != "double")
  remits$isCTR <- as.double(remits$long)
if (typeof(remits$IND_remit) != "integer" && typeof(remits$IND_remit) != "double" )
  remits$IND_remit <- as.double(remits$IND_remit)

typeof(remits$isCTR)

```

## 02: Plot test

```{r}
# coordinate validation
ifelse(max(remits$lat > 90.0) || min(remits$lat < -90.0) || max(remits$long > 180.0) || min(remits$long < -180.0), print("Check your coordinate system..."), print("Coordinates look ok!"))

# test plot
plot(remits$long, remits$lat)
```

## 03: Select origin point from rows

* column "iscenter" value = 1
    + __currently pre-set in test data__
* validate that there is only one center

```{r}
if (sum(remits$isCTR) > 1)
  print("Too many centers. Check data")
if (sum(remits$isCTR) == 0)
  print("No center. Check data. Maybe we'll have an interface to select one later")
if (sum(remits$isCTR) == 1)
  print("There is one center! Hooray!")

ctrX <- remits$long[remits$isCTR == TRUE]
ctrY <- remits$lat[remits$isCTR == TRUE]
ctrPt <- c(ctrX,ctrY)
cat("Your center point has the following coordinates:", ctrPt[1],"longitude and", ctrPt[2], "latitude. ")

```

## 04: Calculate all great circle distances and bearings to center point

* __NB:__ half of circumference of earth is ~20,000 km, which might be half of the max dist we should allow
    + example data max is 7,241 km

```{r}
# create a column in remits, fill with km distances
  remits$distance <- geodist(ctrY, ctrX, remits$lat, remits$long, units = "km")  
  maxdist <- max(remits$distance) # max great circle distance

  print( # print out both the longest axial and great circle distances
    paste0("The longest great circle distance between your center point, ",
           remits$name[remits$isCTR == TRUE], ", and farthest point, ",
           remits$name[remits$distance == max(remits$distance)], ", is ",
           round(max(remits$distance), 0), " km.")
    )
  
for (row in 1:nrow(remits)){ # convert to polar coordinates
  countryLong <- remits[row,"long"]
  countryLat <- remits[row,"lat"]
  countryLonglat <- c(countryLong, countryLat)
  remits[row,"ctrPtGeobearing"] <- bearing(ctrPt, countryLonglat, a=6378137, f=1/298.257223563)
}
  rm(countryLat,countryLong,countryLonglat,row)
  
for (row in 1:nrow(remits)) # convert from geographic bearings to polar coordinates
  {if(remits$ctrPtGeobearing[row] <= 0)
  # if geobearing is 0 or negative, mathbearing is 90 plus bearing
  remits$ctrPtMathbearing[row] <- abs(remits$ctrPtGeobearing[row]) + 90
  if(remits$ctrPtGeobearing[row] > 0 & remits$ctrPtGeobearing[row] < 90)
  # if geobearing is positive and equal to or under 90, mathbearing is 90 - bearing
  remits$ctrPtMathbearing[row] <- (90 - remits$ctrPtGeobearing[row])
  if(remits$ctrPtGeobearing[row] >= 90 & remits$ctrPtGeobearing[row] <= 180)
  # if geobearing is positive and between 90 and 180 (inclusive), mathbearing is 360 - (geobearing - 90)
  remits$ctrPtMathbearing[row] <- 360 - (remits$ctrPtGeobearing[row] - 90)}
  rm(row)

  remits <- select(remits,-starts_with("ctrPtGeo"))
  # later we use the ctrPtMathbearing and the transformed distances to get back to cartesian coordinates

```

## 05: Initial plot
* GGplot of the lat and long with low-res world background

```{r}
# first plot with ctrPt and data visualized, still using lat/long
plot05 <- ggplot(remits, 
       aes(remits$long, 
           remits$lat, 
           color = remits$distance)) + 
           geom_point() + 
           geom_point(data = (as.data.frame(ctrPt)), aes(ctrPt[1], ctrPt[2]), color = "orange")

plot05

```  
  
## 06: Plot 2 - cash input
* Plot witn cash input
  
```{r} 
remits$cashInput <- (remits$IND_remit * 1) + .5 

for (row in 1:nrow(remits)){
  remits[row,"circdistancex"] <- (pol2cart(remits$distance[row], remits$ctrPtMathbearing[row], degrees = TRUE))[1]
  remits[row,"circdistancey"] <- (pol2cart(remits$distance[row], remits$ctrPtMathbearing[row], degrees = TRUE))[2]
}
rm(row)

# Lets make some circles
circles <- data.frame(
  x0 = 0,
  y0 = 0,
  r = seq(1500, 15000, length.out = 7)
)

# Use coord_fixed to ensure true circularity
ggplot() +
  geom_circle(aes(x0 = x0, y0 = y0, r = r), colour = "orange", data = circles, show.legend = NA, inherit.aes = FALSE) +
  coord_fixed()

plot06 <- ggplot(remits, aes(
                remits$circdistancex, 
                remits$circdistancey, 
                color = remits$distance)) +
        geom_circle(aes(x0 = x0, y0 = y0, r = r),
          colour = "orange", 
          data = circles, 
          show.legend = NA, 
          inherit.aes = FALSE) +
        geom_point(stroke = 1, 
                   size = 0.002*(remits$cashInput)
                   ) +
        geom_point(data = (as.data.frame(ctrPt)), 
                   aes(0, 0), 
                   color = "orange") +
        coord_fixed() + 
        labs(color = "Distance in km", x = NULL, y = NULL) +
        guides(colour = "colorbar",size = "legend") +
        theme(panel.background = element_blank())

plot06

```  

## 07: Plot on log scale

```{r} 
# replot on log scale

remits$distancelog <- log(remits$distance)

for (row in 1:nrow(remits)){
  remits[row,"distancelogcircx"] <- (pol2cart(remits$distancelog[row], remits$ctrPtMathbearing[row], degrees = TRUE))[1]
  remits[row,"distancelogcircy"] <- (pol2cart(remits$distancelog[row], remits$ctrPtMathbearing[row], degrees = TRUE))[2]
}

logcircles <- data.frame()

for (row in 1:nrow(circles)){
  logcircles[row,"r"] <- log(circles[row,"r"])
  logcircles[row,"x0"] <- circles[row,"x0"]
  logcircles[row,"y0"] <- circles[row,"y0"]
}
rm(row)

plot07 <- ggplot(remits, aes(
                remits$distancelogcircx, 
                remits$distancelogcircy, 
                color = remits$distance)) + 
  geom_circle(aes(x0 = x0, y0 = y0, r = r), 
                      colour = "orange", data = logcircles, 
                      show.legend = NA, inherit.aes = FALSE) +
  geom_point(stroke = 1, 
             size = 0.002*(remits$cashInput)) + 
  geom_point(data = (as.data.frame(ctrPt)), 
             aes(0, 0), 
             color = "orange") +
  coord_fixed() + 
  labs(color = "Great circle distance in km", x = NULL, y = NULL) +
  theme(panel.background = element_blank(), 
        axis.ticks = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank())

plot07

```  

## 08 Plot on square root scale (with and without scale circles)

```{r} 
remits$distancesqrt <- sqrt(remits$distance)

for (row in 1:nrow(remits)){
  remits[row,"sqrtcircX"] <- (pol2cart(remits$distancesqrt[row], remits$ctrPtMathbearing[row], degrees = TRUE))[1]
  remits[row,"sqrtcircY"] <- (pol2cart(remits$distancesqrt[row], remits$ctrPtMathbearing[row], degrees = TRUE))[2]
}
rm(row)

plot08A <- ggplot(remits, aes( # plot with scale circles
                remits$sqrtcircX, 
                remits$sqrtcircY, 
                color = remits$distance)) +  
          geom_circle(aes(x0 = x0, y0 = y0, r = r), 
                      colour = "orange", data = sqrt(circles), 
                      show.legend = NA, inherit.aes = FALSE) + 
          geom_point(stroke = 1, 
                     size = 0.002*(remits$cashInput)) + 
          geom_point(data = (as.data.frame(ctrPt)), aes(0, 0), color = "orange") + 
          coord_fixed() + 
          labs(color = "Distance in km", x = NULL, y = NULL) + 
          theme(panel.background = element_blank(), axis.ticks = element_blank(),
                axis.text.x = element_blank(),axis.text.y = element_blank())

plot08A

plot08B <- ggplot(remits, aes( # plot without circles
                remits$sqrtcircX, 
                remits$sqrtcircY, 
                color = remits$distance)) + 
          geom_point(stroke = 1, 
                     size = 0.002*(remits$cashInput)) + 
          geom_point(data = (as.data.frame(ctrPt)), 
                     aes(0, 0), color = "orange") + 
          coord_fixed() + 
          labs(color = "Distance in km", x = NULL, y = NULL) + 
          theme(panel.background = element_blank(), axis.ticks = element_blank(),
                axis.text.x = element_blank(),axis.text.y = element_blank())

plot08B

```

## 09: Manual Control Points
* Define a decay function using manual control points

```{r}
# polynom::poly
# code to find a Lagrange interpolation polynomial for our control points. for now I'm just defining some, later we should explore the user input and/or data-driven methods.

geogdist <- c(0, 4000, 8000, maxdist) # let's say under 500 km is close, over 2000km is far
chartdist <- c(0, 400, 800, 1200) # this just gets us equal intervals on the graph for the different segments of the lines

dat <- data.frame(cbind(geogdist, chartdist)) # chart points to check intervals
ggplot(dat, aes(x=geogdist, y=chartdist)) + geom_point(size=5, col='blue') 

lagrangepoly <- poly.calc(geogdist, chartdist) # calculate polynomial from control points
lagrangepoly # print polynomial 

curve(predict(lagrangepoly,x), from = 0, to = maxdist, n = 100)

predict(lagrangepoly, 200) # print predicted value for geogdist of 200

remits$lagrangedistcirc <- predict(lagrangepoly, remits$distance) # set new lagrange great circle distances using new polynomial

# function to make new circles with any lagrange polynomial set above
lagrange_predict <- function(circlesdataframe) { 
  lagrangecirclesdataframe <- circlesdataframe
  for (row in 1:nrow(circlesdataframe)){
    lagrangecirclesdataframe[row,"r"]<- predict(lagrangepoly, circlesdataframe[row,"r"])}
  return(lagrangecirclesdataframe)
}

lagrangecircles <- lagrange_predict(circles) # check if it worked
lagrangecircles

```

## 10: Plot with manual control points (with or without scale circles)

```{r}

for (row in 1:nrow(remits)){ # calculate new coordinates using lagrange distances
  remits[row,"lagrangedistcircX"] <- (pol2cart(remits$lagrangedistcirc[row], remits$ctrPtMathbearing[row], degrees = TRUE))[1] # returns x coordinate only
  remits[row,"lagrangedistcircY"] <- (pol2cart(remits$lagrangedistcirc[row], remits$ctrPtMathbearing[row], degrees = TRUE))[2] # returns y coordinate only
}
rm(row)

plot10A <- ggplot(remits, aes( # plot with 500-km circles
                lagrangedistcircX, 
                lagrangedistcircY, 
                color = remits$distance)) +  
          geom_circle(aes(x0 = x0, y0 = y0, r = r), 
                      colour = "orange", data = lagrange_predict(circles), 
                      show.legend = NA, inherit.aes = FALSE) + 
          geom_point(stroke = 1, 
                     size = 0.002*(remits$cashInput)) + 
          geom_point(data = (as.data.frame(ctrPt)), aes(0, 0), color = "orange") + 
          coord_fixed() + 
          labs(color = "Distance in km", x = NULL, y = NULL) + 
          theme(panel.background = element_blank(), 
                axis.ticks = element_blank(), 
                axis.text.x = element_blank(),
                axis.text.y = element_blank())
plot10A

plot10B <- ggplot(remits, aes( # plot without circles, zoomed in on data (do we need x and y lims again to do this programmatically?)
                lagrangedistcircX, 
                lagrangedistcircY, 
                color = remits$distance)) + 
          geom_point(stroke = 1, 
                     size = 0.002*(remits$cashInput)) + 
          geom_point(data = (as.data.frame(ctrPt)), aes(0, 0), color = "orange") + 
          coord_fixed() + 
          labs(color = "Distance in km", x = NULL, y = NULL) + 
          theme(panel.background = element_blank(), 
                axis.ticks = element_blank(), 
                axis.text.x = element_blank(),
                axis.text.y = element_blank())
plot10B

```

# 11 Interpolation & Integrated plotting

```{r}

csh <- NULL
csh <- with(remits, interp(x = lagrangedistcircX, y = lagrangedistcircY, z = IND_remit))

# Base R version
plot11A <- filled.contour(x = csh$x,
               y = csh$y,
               z = csh$z,
               color.palette = colorRampPalette(c("white", "blue")),
               main = "Remittances to India") +
               xlim(-2000,2000) + 
               ylim(-2000,2000)
plot11A

#########

# ggplot version

csh_grid_df <- melt(csh$z, na.rm = TRUE)
names(csh_grid_df) <- c("x", "y", "cash")
csh_grid_df$ldcX <- csh$x[csh_grid_df$x]
csh_grid_df$ldcY <- csh$y[csh_grid_df$y]

plot11B <- ggplot(data = csh_grid_df, 
                 aes(x = ldcX, 
                     y = ldcY, 
                     z = cash)) +
          geom_tile(aes(fill = cash)) +
          stat_contour(bins = 15, 
                       color = "black", 
                       size = 0.1) +
          scale_fill_gradient(name = "cash", 
                              low = "yellow", 
                              high = "red")

plot11B

############################################  
############################################

# Integrated plots, incl sec 10

remitsSub <- remits[!(remits$isCTR == TRUE), ]
# holy moly this is a sloppy solution, but it will have to do for now

plot11C <- ggplot() +
          geom_circle(aes(x0 = x0, y0 = y0, r = r), 
                      colour = "#f7f7f7", 
                      data = lagrange_predict(circles),
                      size = 3,
                      show.legend = NA, 
                      inherit.aes = FALSE) +        
          geom_contour(data = csh_grid_df, 
                 aes(x = ldcX, 
                     y = ldcY, 
                     z = csh_grid_df$cash,
                     color = stat(level)),
                 bins = 15,  
                 size = 0.5) +
          scale_colour_gradient(low = "cornflowerblue", high = "red") +
  ### begin simulated drop shadow
            geom_text(data = remitsSub,
                    aes(lagrangedistcircX,
                        lagrangedistcircY,
                        label= remitsSub$name),
                    size = 1.8,
                    check_overlap = TRUE,
                    color = "white") +
            geom_text(data = remitsSub,
                    aes(lagrangedistcircX,
                        lagrangedistcircY,
                        label= remitsSub$name),
                    size = 2,
                    check_overlap = TRUE,
                    color = "white") +
  ### end simulated drop shadow
            geom_text(data = remitsSub,
                    aes(lagrangedistcircX,
                        lagrangedistcircY,
                        label= remitsSub$name,
                        color = remitsSub$cashInput),
                    size = 1.5,
                    check_overlap = TRUE) +
            geom_point(data = (as.data.frame(ctrPt)), 
                     aes(ctrPt[1], ctrPt[2]), 
                     color = "black", 
                     shape = 8) +
         geom_text(data = (as.data.frame(ctrPt)), 
                   geom = "text", 
                   x = ctrPt[1], y = ctrPt[2], 
                   label = remits$name[remits$isCTR == TRUE],
                   color = "black",
                   size = 2,
                   vjust = 1.25,
                   hjust = -0.25) +
          theme(panel.background = element_blank(), 
                axis.ticks = element_blank(), 
                axis.text.x = element_blank(),
                axis.text.y = element_blank(),
                axis.title = element_blank(),
                legend.position = 'none')
         # scale_fill_gradient(name = "cash", low = "yellow", high = "red")

plot11C


plot11D <- ggplot() +
          geom_circle(aes(x0 = x0, y0 = y0, r = r), 
                      colour = "#f7f7f7", 
                      data = lagrange_predict(circles),
                      size = 3,
                      show.legend = NA, 
                      inherit.aes = FALSE) +        
          geom_contour(data = csh_grid_df, 
                 aes(x = ldcX, 
                     y = ldcY, 
                     z = csh_grid_df$cash,
                     color = stat(level)),
                 bins = 15,  
                 size = 0.5) +
          scale_colour_gradient(low = "cornflowerblue", high = "red") +
          geom_point(data = remitsSub,
                aes(lagrangedistcircX,
                    lagrangedistcircY,
                    label= remitsSub$name,
                    alpha = -(remitsSub$lagrangedistcirc)),
                size = .3,
                color="gray35") +
          geom_point(data = (as.data.frame(ctrPt)), 
                     aes(ctrPt[1], ctrPt[2]), 
                     color = "orange", 
                     shape = 8) +
         geom_text(data = (as.data.frame(ctrPt)), 
                   geom = "text", 
                   x = ctrPt[1], y = ctrPt[2], 
                   label = remits$name[remits$isCTR == TRUE],
                   color = "orange",
                   size = 2,
                   vjust = 1.25,
                   hjust = -0.25) +
          theme(panel.background = element_blank(), 
                axis.ticks = element_blank(), 
                axis.text.x = element_blank(),
                axis.text.y = element_blank(),
                axis.title = element_blank(),
                legend.position = 'none')
         # scale_fill_gradient(name = "cash", low = "yellow", high = "red")

plot11D

```

# 12 New Lagrange Methods Experiments

```{r} 

dat <- data.frame(cbind(geogdist, chartdist)) # chart points to check intervals
lagrangemonpol <- monpol(formula = chartdist~geogdist, data = dat, plot.it = TRUE,
                         ptype = "SOS", monotone = "increasing")

lagrangemonpol
predict(lagrangemonpol,100)

predict(lagrangemonpol, c(0,100,200,300,400,500,600), scale = c("original"))

curve(predict(lagrangemonpol,x), from = 0, to = maxdist, n = 4)

print(lagrangemonpol, digits = max(3, getOption("digits") - 3))

# plot(lagrangemonpol) #EM 3/29: this breaks for me. "'x' is a list, but does not have components 'x' and 'y'"

median(remits[["distance"]])
geogdist <- c(0, 1220.595, maxdist) # let's say under 500 km is close, over 2000km is far
chartdist <- c(0, 600, 1200) # this just gets us equal intervals on the graph for the different segments of the lines

# abline(lm(chartdist~geogdist)) #EM 3/29: breaks for me, possibly due to line 450

```


# 13 

```{r}

```

