---
title: "PCP_betaWorkflow"
author: "Will P."
date: "5/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 00: R Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(sp)
library(ggplot2)
library(ggforce) # to plot circles
library(dplyr)
library(tidyverse)
library(geosphere) # for geodist
library(useful) # for cartesian conversions
library(akima) # for irregular grid
library(reshape2) # for contor/grid


```

## 01: Input CSV into R
* Read and give headers and datatypes for each (eventually do more sophisticated validation/type detection)

```{r}
# TEST DATA: only one at a time
dataFile <- "testData/IND_remittances.csv" # remittances to India in USD in 201X
# dataFile <- "testData/routes_bare_sm.csv" # flights to DTW
# dataFile <- "testData/routes_sm.csv" # flights to DTW with ORD and MSP options
# dataFile <- "testData/china1911_XJofficials.csv" # officials from Xinjiang in 1911

df <- read.csv(file = dataFile)

summary(df)
class(df) # test to make sure it's a data frame

# STORE COL NAMES FOR LATER

colListOrig <- colnames(df)
latNames <- list("lat", "latitude", "y")
lonNames <- list("lon", "long", "longitude", "x")

# STARTING TO INCLUDE FORCED GENERIC COLUMN HEADINGS

#for loop
    for (col in 1:ncol(df)){ 
    # lat
    (max(routes$lat <= 90.0) || min(routes$lat >= -90.0)
    # long
    max(routes$long <= 180.0) || min(routes$long >= -180.0)
    # ctrBin
    (min = 0 && max = 1 && sum = 1) or typeof == "logical"
    # valName
    typeof == "character"
    # val
    typeof == "integer" || "double"

typeof(routes$isDTW)

     # obnoxiously explicit validation

if (typeof(routes$airport) != "character")
  routes$airport <- as.character(routes$airport)
if (typeof(routes$isDTW) != "logical")
  routes$isDTW <- as.logical(routes$isDTW)
if (typeof(routes$lat) != "double")
  routes$isDTW <- as.double(routes$lat)
if (typeof(routes$long) != "double")
  routes$isDTW <- as.double(routes$long)
if (typeof(routes$flightsDTW) != "integer" && typeof(routes$flightsDTW) != "double" )
  routes$flightsDTW <- as.double(routes$flightsDTW)

typeof(routes$isDTW)

```



## 02: Plot test

```{r}

# test plot
plot(df$long, df$lat)

#temp
ctrX <- df$long[df$isCTR == TRUE]
ctrY <- df$lat[df$isCTR == TRUE]

#real
# ctrX <- df$long[df$ctrBin == TRUE]
# ctrY <- df$lat[df$ctrBin == TRUE]
ctrPt <- c(ctrX,ctrY)
cat("Your center point has the following coordinates:", ctrPt[1],"longitude and", ctrPt[2], "latitude. ")
```




## 04: Calculate all great circle distances and bearings to center point

* __NB:__ half of circumference of earth is ~20,000 km, which might be half of the max dist we should allow
    + example data max is 7,241 km

```{r}
# create a column in df, fill with km distances
  df$distance <- geodist(ctrY, ctrX, df$lat, df$long, units = "km")  
  maxdist <- max(df$distance) # max great circle distance

  print( # print out both the longest axial and great circle distances
    paste0("The longest great circle distance between your center point, ",
           df$valName[df$ctrBin == TRUE], ", and farthest point, ",
           df$valName[df$distance == max(df$distance)], ", is ",
           round(max(df$distance), 0), " km.")
    )
  
for (row in 1:nrow(df)){ # convert to polar coordinates
  valNamelong <- df[row,"long"]
  valNamelat <- df[row,"lat"]
  valNamelonglat <- c(valNamelong, valNamelat)
  df[row,"ctrPtGeobearing"] <- bearing(ctrPt, valNamelonglat, a=6378137, f=1/298.257223563)
}
  rm(valNamelat,valNamelong,valNamelonglat,row)
  
for (row in 1:nrow(df)) # convert from geographic bearings to polar coordinates
  {if(df$ctrPtGeobearing[row] <= 0)
  # if geobearing is 0 or negative, mathbearing is 90 plus bearing
  df$ctrPtMathbearing[row] <- abs(df$ctrPtGeobearing[row]) + 90
  if(df$ctrPtGeobearing[row] > 0 & df$ctrPtGeobearing[row] < 90)
  # if geobearing is positive and equal to or under 90, mathbearing is 90 - bearing
  df$ctrPtMathbearing[row] <- (90 - df$ctrPtGeobearing[row])
  if(df$ctrPtGeobearing[row] >= 90 & df$ctrPtGeobearing[row] <= 180)
  # if geobearing is positive and between 90 and 180 (inclusive), mathbearing is 360 - (geobearing - 90)
  df$ctrPtMathbearing[row] <- 360 - (df$ctrPtGeobearing[row] - 90)}
  rm(row)

  df <- select(df,-starts_with("ctrPtGeo"))
  # later we use the ctrPtMathbearing and the transformed distances to get back to cartesian coordinates

```

## 05: Initial plot
* GGplot of the lat and long with low-res world background

```{r}
# first plot with ctrPt and data visualized, still using lat/long
plot05 <- ggplot(df, 
       aes(df$long, 
           df$lat, 
           color = df$distance)) + 
           geom_point() + 
           geom_point(data = (as.data.frame(ctrPt)), aes(ctrPt[1], ctrPt[2]), color = "orange")

plot05

```  

## 06: Plot 2 - val center connectivity
* Plot connectivity
  
```{r} 
#temp
df$valTrans <- (df$IND_remit * 1) + .5 
#real
#df$valTrans <- (df$val * 1) + .5 

for (row in 1:nrow(df)){
  df[row,"circdistancex"] <- (pol2cart(df$distance[row], df$ctrPtMathbearing[row], degrees = TRUE))[1]
  df[row,"circdistancey"] <- (pol2cart(df$distance[row], df$ctrPtMathbearing[row], degrees = TRUE))[2]
}
rm(row)

# Lets make some circles
circles <- data.frame(
  x0 = 0,
  y0 = 0,
  r = seq(0, maxdist,length.out = 10)
)

# Use coord_fixed to ensure true circularity
ggplot() +
  geom_circle(aes(x0 = x0, y0 = y0, r = r), colour = "orange", data = circles, show.legend = NA, inherit.aes = FALSE) +
  coord_fixed()

plot06 <- ggplot(df, aes(
                df$circdistancex, 
                df$circdistancey, 
                color = df$distance)) +
        geom_circle(aes(x0 = x0, y0 = y0, r = r),
          colour = "orange", 
          data = circles, 
          show.legend = NA, 
          inherit.aes = FALSE) +
        geom_point(stroke = 1, size = df$valTrans) +
        geom_point(data = (as.data.frame(ctrPt)), aes(0, 0), color = "orange") +
        coord_fixed() + labs(color = "Distance in km", x = NULL, y = NULL) +
        guides(colour = "colorbar",size = "legend") +
        theme(panel.background = element_blank())

plot06

```  

## 07: Plot on log scale
an option we can give our users later
```{r} 
# replot on log scale

df$distancelog <- log(df$distance)

for (row in 1:nrow(df)){
  df[row,"distancelogcircx"] <- (pol2cart(df$distancelog[row], df$ctrPtMathbearing[row], degrees = TRUE))[1]
  df[row,"distancelogcircy"] <- (pol2cart(df$distancelog[row], df$ctrPtMathbearing[row], degrees = TRUE))[2]
}

logcircles <- data.frame()

for (row in 1:nrow(circles)){
  logcircles[row,"r"] <- log(circles[row,"r"])
  logcircles[row,"x0"] <- circles[row,"x0"]
  logcircles[row,"y0"] <- circles[row,"y0"]
}
rm(row)

plot07 <- ggplot(df, aes(
                df$distancelogcircx, 
                df$distancelogcircy, 
                color = df$distance)) + 
  geom_circle(aes(x0 = x0, y0 = y0, r = r), 
                      colour = "orange", data = logcircles, 
                      show.legend = NA, inherit.aes = FALSE) +
  geom_point(stroke = 1, size = df$valTrans) + 
  geom_point(data = (as.data.frame(ctrPt)), aes(0, 0), color = "orange") +
  coord_fixed() + labs(color = "Great circle distance in km", x = NULL, y = NULL) +
  theme(panel.background = element_blank(), axis.ticks = element_blank(),
                axis.text.x = element_blank(),axis.text.y = element_blank())

plot07

```  

## 08 Plot on square root scale (with and without scale circles)

```{r} 
df$distancesqrt <- sqrt(df$distance)

## *** GENERALIZE pol2cart function so can run on X rows and don't have to re-deploy every time 
for (row in 1:nrow(df)){ 
  df[row,"sqrtcircX"] <- (pol2cart(df$distancesqrt[row], df$ctrPtMathbearing[row], degrees = TRUE))[1]
  df[row,"sqrtcircY"] <- (pol2cart(df$distancesqrt[row], df$ctrPtMathbearing[row], degrees = TRUE))[2]
}
rm(row)

plot08A <- ggplot(df, aes( # plot with scale circles
                df$sqrtcircX, 
                df$sqrtcircY, 
                color = df$distance)) +  
          geom_circle(aes(x0 = x0, y0 = y0, r = r), 
                      colour = "orange", data = sqrt(circles), 
                      show.legend = NA, inherit.aes = FALSE) + 
          geom_point(stroke = 1, size = df$valTrans) + 
          geom_point(data = (as.data.frame(ctrPt)), aes(0, 0), color = "orange") + 
          coord_fixed() + 
          labs(color = "Distance in km", x = NULL, y = NULL) + 
          theme(panel.background = element_blank(), axis.ticks = element_blank(),
                axis.text.x = element_blank(),axis.text.y = element_blank())

plot08A

plot08B <- ggplot(df, aes( # plot without circles
                df$sqrtcircX, 
                df$sqrtcircY, 
                color = df$distance)) + 
          geom_point(stroke = 1, size = df$valTrans) + 
          geom_point(data = (as.data.frame(ctrPt)), aes(0, 0), color = "orange") + 
          coord_fixed() + 
          labs(color = "Distance in km", x = NULL, y = NULL) + 
          theme(panel.background = element_blank(), axis.ticks = element_blank(),
                axis.text.x = element_blank(),axis.text.y = element_blank())

plot08B

```


