---
title: "PCP_betaWorkflow"
author: "Will P."
date: "5/30/2019"
output: html_document
---

## 00: R Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(sf)
library(sp) # might not need
library(ggplot2)
library(ggforce) # to plot circles
library(dplyr)
library(tidyverse)
library(tools) #to learn file extension
library(jsonlite) # for loading in json files
library(geojsonio) # for loading geojson
# 
library(geosphere) # where we get bearing
library(gmt) # actually for geodist
library(useful) # for cartesian conversions

# NB: to check later as we re-visit new contours
library(akima) # for irregular grid
library(reshape2) # for contor/grid
library(contoureR) # for contours w/o grid
library(deldir) # CHECK: trying for TIN generation, see sec. 11
library(lawn) # CHECK:  trying for TIN generation, see sec. 11


# FUNCTIONS WE'RE USING
#source("PCP_functions/convertToPolar.R")

```

## 01: Input CSV into R
* Read and give headers and datatypes for each (eventually do more sophisticated validation/type detection)

```{r}
# TEST DATA: only one at a time
# dataFile <- "testData/IND_remittances.csv" # remittances to India in USD in 201X
# dataFile <- "testData/trampCities.csv" # trampoline city data
# dataFile <- "testData/trampZips.csv" # trampoline zipcode data
# dataFile <- "testData/routes_bare_sm.csv" # flights to DTW
# dataFile <- "testData/routes_sm.csv" # flights to DTW with ORD and MSP options
# dataFile <- "testData/china1911_XJofficials.csv" # officials from Xinjiang in 1911
dataFile <- "testData/Zagat1991Geocoded.geojson" # Will's geocoded Zagat data--mostly just for geojson testing


# Decide file type
df_ext <- file_ext(dataFile)

if(df_ext == "csv"){
  df <- read.csv(file = dataFile)
  print("hooray, a CSV!")
} else {
  if(df_ext == "json"){
    df <- as.data.frame(read_json(dataFile, simplifyVector = TRUE))
    print(paste("this is a json, right?"))
    # NB: we may need to do more to this
    # strip features.properties from each column heading
  } else {
    if(df_ext == "geojson"){
      df <- st_as_sf(geojson_read(dataFile, method = "local", parse = TRUE, what = "sp"))
      print("this is a geojson! yeah!")
    } else {
      print("what kind of file even is this?")
    }
  }
}
   
  

summary(df)
class(df[["geometry"]]) # test to make sure it's a data frame

colListOrig <- colnames(df) # store column names for later
latNames <- list("lat","Lat","LAT", "latitude", "Latitude", "LATITUDE", "y","Y", "coords.x2") # add as they come up
lonNames <- list("lon","Lon","LON","long","Long","LONG","longitude", "Longitude", "LONGITUDE", "x","X", "coords.x1")
valNameChoices <- list()
valChoices <- list()

df2 <- df # cloning df for non-destructive editing

latflag <- 0 # need these here for the column detection
lonflag <- 0
ctrBinflag <- 0
valNameflag <- 0
valflag <- 0

for (col in 1:ncol(df)) {
  if (typeof(df[[col]]) == "double"
  && latflag == 0
  && max(as.numeric(df[[col]]), na.rm = T) <= 90.0
  && min(as.numeric(df[[col]]), na.rm = T) >= -90.0
  && names(df)[[col]] %in% latNames) # lat
  { print(paste("Found a lat column: ", names(df)[[col]]))
    df2$lat <- df[[col]]
    latflag <- 1}
  else{
    if (typeof(df[[col]]) == "double"
    && lonflag == 0
    && max(as.numeric(df[[col]]), na.rm = T) <= 180.0
    && min(as.numeric(df[[col]]), na.rm = T) >= -180.0
    && names(df)[col] %in% lonNames) # lon
    { print(paste("Found a lon column: ", names(df)[[col]]))
      df2$lon <- df[[col]]
      lonflag <- 1}
    else{
      if (typeof(df[[col]]) != "character"
      && typeof(df[[col]]) != "list"
      && min(as.numeric(df[[col]]), na.rm = T) == 0
      && max(as.numeric(df[[col]]), na.rm = T) == 1
      && sum(as.numeric(df[[col]]), na.rm = T) == 1) # ctrBin
      { df2$ctrBin <- as.logical(df[[col]])
        print(paste("Found a ctrBin column: ", names(df)[[col]]))
        ctrBinflag <- 1}
      else{
        if (typeof(df[[col]]) == "character" # catches name and name_long
        || is.factor(df[[col]]) == T) # valName
        { if (valNameflag == 0)
            {df2$valName <- as.character(df[[col]])
            print(paste("Found a valName column: ", names(df)[[col]]))
            valNameflag <- 1}
            else{
              print(paste("Found an alternate valName column: ", names(df)[[col]]))
              valNameChoices <- c(valNameChoices, names(df)[[col]])}
              # NOTE this isn't storing anywhere yet, might need to later for input
          }
        else{
          if (typeof(df[[col]]) == "integer" || typeof(df[[col]]) == "double" # val
          && !(names(df)[col] %in% lonNames)
          && !(names(df)[col] %in% latNames)
          && names(df[col]) != "geometry")
          { if (valflag == 0)
            {print(paste("Think it's a val!"))
            df2$val <- as.double(df[[col]])
            print(paste("Found a val column: ", names(df)[[col]]))
            valflag <- 1}
            else{
              print(paste("Found an alternate val column: ", names(df)[[col]]))
              valChoices <- c(valChoices, names(df)[[col]])}
              # NB: NOTE this isn't storing anywhere yet, might need to later for input
          }
        }
      }
    }
  }
}

if("ctrBin" %in% names(df)[col]) {
  print("you prepared your data soooooo well!")
} else {
  df2$ctrBin <- FALSE
}

if(df_ext == "csv"){
  df2 <- dplyr::select(df2, valName, val, lat, lon, ctrBin) # just the fields we want
} else {
  if(df_ext == "geojson"){
    df2 <- dplyr::select(df2, valName, val, lat, lon, ctrBin) #something else 
  } else {
    print(paste("what even is this?"))
  }
}



```

## 02: Plot Test

```{r}

# Test plot
plot(df2$lon, df2$lat)

# Store and print center point
ctrPtName <- df2$valName[df2$ctrBin == TRUE]
ctrPt <- c(df2$lat[df2$ctrBin == TRUE],df2$lon[df2$ctrBin == TRUE]) # get lat/lon
cat("Your center point, ",ctrPtName,", is located at", ctrPt[1],"longitude and", ctrPt[2], "latitude. ")

```

## 03: Calculate all great circle distances and bearings to center point

```{r}
# create a column in df, fill with km distances
  df2$distance <- geodist(ctrPt[1], ctrPt[2], df2$lat, df2$lon, units = "km")  
  maxdist <- max(df2$distance) # max great circle distance
  
  print( # print out both the longest axial and great circle distances
    paste0("The great circle distance between your center point, ",
           ctrPtName, ", and farthest point, ",
           df2$valName[df2$distance == max(df2$distance)], ", is ",
           round(maxdist, 0), " km.")
    )
  if(maxdist > 20000){
    print("Your maximum distance is more than half of the Earth's circumference; things might get a little squirrely!")
    } else {
    print("Your maximum distance is less than half of the Earth's circumference! Nice!")
    }
  
  df2 <- df2 %>% mutate(ctrPtGeobearing = geosphere::bearing(c(ctrPt[2],ctrPt[1]), cbind(lon, lat), a=6378137, f=1/298.257223563)) # get bearing of all points to center 

  # THIS NEXT PART SHOULD BE A FUNCTION TOO
  
for (row in 1:nrow(df2)) # convert from geographic bearings to polar coordinates
  {if(df2$ctrPtGeobearing[row] <= 0)
  # if geobearing is 0 or negative, mathbearing is 90 plus bearing
  df2$ctrPtMathbearing[row] <- abs(df2$ctrPtGeobearing[row]) + 90
  if(df2$ctrPtGeobearing[row] > 0 & df2$ctrPtGeobearing[row] < 90)
  # if geobearing is positive and equal to or under 90, mathbearing is 90 - bearing
  df2$ctrPtMathbearing[row] <- (90 - df2$ctrPtGeobearing[row])
  if(df2$ctrPtGeobearing[row] >= 90 & df2$ctrPtGeobearing[row] <= 180)
  # if geobearing is positive and between 90 and 180 (inclusive), mathbearing is 360 - (geobearing - 90)
  df2$ctrPtMathbearing[row] <- 360 - (df2$ctrPtGeobearing[row] - 90)}
  rm(row)

  df2 <- dplyr::select(df2,-starts_with("ctrPtGeo"))
  # later we use the ctrPtMathbearing and the transformed distances to get back to cartesian coordinates

```

## 04: Initial plot
* GGplot of the lat and long with low-res world background (NOTE WHERE IS THE BACKGROUND?)

```{r}
# first plot with ctrPt and data visualized, still using lat/long

# df2 <- subset(df2,df2$val > 49) # FOR TESTING try again with a subset
maxdist <- max(df2$distance) # max great circle distance

plot05 <- ggplot(df2, 
       aes(df2$lon, 
           df2$lat, 
           color = df2$distance)) + 
           geom_point() + 
           geom_point(data = (as.data.frame(ctrPt)), aes(ctrPt[2], ctrPt[1]), color = "orange")

plot05

```  

## 05: Transform data values for effective visualization later

```{r} 
#temp ***
df2$valTrans <- log(df2$val) + 1 # manual one I did for NH data to see what we're looking for
#df2$valTrans <- (df2$val / 400)^1.5 # manual one I did for NH data to see what we're looking for

hist(df2$val, breaks = 20) # just for giggles
hist(sqrt(df2$val), breaks = 20)
hist(log(df2$val), breaks = 20)

#df2$val[df2$val == 0] <- NA # turn zeros to NAs for our purposes

print(min(df2$val, na.rm = T)) # find lowest non-zero value (that will be radius 1)
print(max(df2$val, na.rm = T)) # find highest non-zero value (that will be radius X)
print(median(df2$val, na.rm = T)) # find median value (that could be radius 1 + X / 2)

#min(my_data_frame[my_data_frame$my_column_number>0,my_column_number])

# *** NEED DATA DRIVEN TRANSFORMATION, not best guess from old test data
#df2$valTrans <- (df2$val * 1) + .5 # MANUAL FOR AIRPORT DATA

```  

## 06: Plot Great Circle Distances

```{r} 

# Run polar to cartesian conversion on x and y axes
df2 <- df2 %>% mutate(circdistancex =  (useful::pol2cart(distance,ctrPtMathbearing,degrees = TRUE)[[1]]))
df2 <- df2 %>% mutate(circdistancey = (useful::pol2cart(distance,ctrPtMathbearing,degrees = TRUE)[[2]]))

# Combine x and y into a matrix, add as a column, remove x and y columns
df2$circcoords <- cbind(df2$circdistancex,df2$circdistancey)
df2 <- dplyr::select(df2,-starts_with("circdistance"))

# Lets make some circles starting at the center point going to maxdist
circles <- data.frame(
  x0 = 0,
  y0 = 0,
  r = seq(0, maxdist,length.out = 11)
)
circles <- circles[-1,] # remove zero-radius circle to help with log transformation

# Use coord_fixed to ensure true circularity

plot06 <- ggplot(df2, aes(
                df2$circcoords[,1], 
                df2$circcoords[,2], 
                color = df2$distance)) +
        geom_circle(aes(x0 = x0, y0 = y0, r = r),
          colour = "orange", 
          data = circles, 
          show.legend = NA, 
          inherit.aes = FALSE) +
        geom_point(stroke = 1, size = df2$valTrans) +
        geom_point(data = (as.data.frame(ctrPt)), aes(0, 0), color = "orange") +
        coord_fixed() + labs(color = paste0("Distance from ",ctrPtName," (km)"), x = NULL, y = NULL) +
        guides(colour = "colorbar",size = "legend") +
        theme(panel.background = element_blank())

plot06

```  

## 07: Plot on Log Scale
An option we can give our users later

```{r} 
# Replot coordinates on log distance scale
df2 <- df2 %>% mutate(
  logdistancex =  (useful::pol2cart(log(distance),ctrPtMathbearing,degrees = TRUE)[[1]]), 
  logdistancey = (useful::pol2cart(log(distance),ctrPtMathbearing,degrees = TRUE)[[2]])
  )

# Combine x and y into a matrix, add as a column, remove x and y columns
df2$logcoords <- cbind(df2$logdistancex,df2$logdistancey)
df2 <- dplyr::select(df2,-starts_with("logdistance"))

# NEED TO OVERRIDE CENTER POINT INFINITE VALUES WITH 0,0

# Plot it

plot07 <- ggplot(df2, aes(
                df2$logcoords[,1], 
                df2$logcoords[,2], 
                color = df2$distance)) + 
  geom_circle(aes(x0 = x0, y0 = y0, r = log(r)), 
                      colour = "orange", data = circles, 
                      show.legend = NA, inherit.aes = FALSE) +
  geom_point(stroke = 1, size = df2$valTrans) + 
  geom_point(data = (as.data.frame(ctrPt)), aes(0, 0), color = "orange") +
  coord_fixed() + labs(color = paste0("Distance from ",ctrPtName," (km)"), x = NULL, y = NULL) +
  theme(panel.background = element_blank(), axis.ticks = element_blank(),
                axis.text.x = element_blank(),axis.text.y = element_blank())

plot07

df2 <- dplyr::arrange(df2, -val) # sorting for draw order below
df2$num <- ave(df2$val, FUN = seq_along) # also sorting?

plot07B <-  ggplot(df2 %>%
                     arrange(desc(val)),
                     aes(
                df2$logcoords[,1], 
                df2$logcoords[,2], 
                color = df2$val,
                order=df2$num)) + 
  geom_circle(aes(x0 = x0, y0 = y0, r = log(r)), 
                      colour = "grey65", data = circles, 
                      show.legend = NA, inherit.aes = FALSE) +
  geom_point(stroke = 1, size = df2$valTrans) + 
  geom_text(data = df2,
                    aes(df2$logcoords[,1],
                        df2$logcoords[,2],
                        label= df2$valName),
                    size = 3,
                    check_overlap = TRUE,
                    color = "White") +
  coord_fixed() + labs(color = paste0("Visits to ",ctrPtName), x = NULL, y = NULL) +
  theme(panel.background = element_rect(fill = "grey50", linetype = "blank"), axis.ticks = element_blank(),
                axis.text.x = element_blank(),axis.text.y = element_blank(), panel.grid = element_blank())

plot07B




```  

## 08 Plot on Square Root Scale

```{r} 

# Replot coordinates on square root distance scale
df2 <- df2 %>% mutate(
  sqrtdistancex =  (useful::pol2cart(sqrt(distance),ctrPtMathbearing,degrees = TRUE)[[1]]), 
  sqrtdistancey = (useful::pol2cart(sqrt(distance),ctrPtMathbearing,degrees = TRUE)[[2]])
  )

# Combine x and y into a matrix, add as a column, remove x and y columns
df2$sqrtcoords <- cbind(df2$sqrtdistancex,df2$sqrtdistancey)
df2 <- select(df2,-starts_with("sqrtdistance"))

df2$sqrtcoords

plot08A <- ggplot(df2, aes( # plot with scale circles
                df2$sqrtcoords[,1], 
                df2$sqrtcoords[,2], 
                color = df2$distance)) +  
          geom_circle(aes(x0 = x0, y0 = y0, r = r), 
                      colour = "orange", data = sqrt(circles), 
                      show.legend = NA, inherit.aes = FALSE) + 
          geom_point(stroke = 1, size = df2$valTrans) + 
          geom_point(data = (as.data.frame(ctrPt)), aes(0, 0), color = "orange") + 
          coord_fixed() + 
          labs(color = paste0("Distance from ",ctrPtName," (km)"), x = NULL, y = NULL) + 
          theme(panel.background = element_blank(), axis.ticks = element_blank(),
                axis.text.x = element_blank(),axis.text.y = element_blank())

plot08A

plot08B <- ggplot(df2, aes( # plot without circles
                df2$sqrtcoords[,1], 
                df2$sqrtcoords[,2], 
                color = df2$distance)) + 
          geom_point(stroke = 1, size = df2$valTrans) + 
          geom_point(data = (as.data.frame(ctrPt)), aes(0, 0), color = "orange") + 
          coord_fixed() + 
          labs(color = paste0("Distance from ",ctrPtName," (km)"), x = NULL, y = NULL) + 
          theme(panel.background = element_blank(), axis.ticks = element_blank(),
                axis.text.x = element_blank(),axis.text.y = element_blank())

plot08B

df2 <- dplyr::arrange(df2, -val)
df2$num <- ave(df2$val, FUN = seq_along)


# df2 <- read.csv(file = "tramp50PlusCities.csv") # TEST with additional variables (have to change things above to to get the circles and other constants right, so this is kind of a mess, don't recommend this way)

plot08C <-  ggplot(df2 %>%
                     arrange(desc(val)),
                     aes(
                df2$sqrtcoords[,1], 
                df2$sqrtcoords[,2], 
                color = df2$valTrans,
                order=df2$num)) + 
  geom_circle(aes(x0 = x0, y0 = y0, r = r), 
                      colour = "grey65", data = sqrt(circles), 
                      show.legend = NA, inherit.aes = FALSE) +
  geom_point(stroke = 1, size = (df2$valTrans^2)/2.5) + 
  geom_text(data = df2,
                    aes(df2$sqrtcoords[,1],
                        df2$sqrtcoords[,2],
                        label= df2$valName),
                    size = 3,
                    check_overlap = FALSE,
                    color = "White") +
  coord_fixed() + 
  labs(color = paste0("Visits/1000 residents"), size = "Size", x = NULL, y = NULL) +
  guides(colour = guide_colourbar(order = 1),
         size = guide_legend(order = 2)) +
  theme(panel.background = element_rect(fill = "grey50", linetype = "blank"), axis.ticks = element_blank(),
                axis.text.x = element_blank(),axis.text.y = element_blank(), panel.grid = element_blank())

plot08C

#write_csv(df2, "tramp50PlusCities.csv")

```

## 09: Come up with Custom Control Points
* Define a distance decay function using data-defined or user-specified control points

```{r}
# polynom::poly
# code to find a stepwise function using our control points.

# Automated route (could also just have user specify neardist and fardist)
hist(df2$distance)
distancemedian <-median(df2$distance)
geogdist <- c(0, distancemedian, ((distancemedian + distancemedian + maxdist)/3), maxdist) # let's say under 500 km is close, over 2000km is far
neardist <- geogdist[2] # these two are the output of geogdist above
fardist <- geogdist[3]
chartdist <- c(0, 400, 800, 1200) # this just gets us equal intervals on the graph for the different segments of the lines

# Simpler Automated route (just median)
hist(df2$distance)
distancemedian <-median(df2$distance)
geogdist <- c(0, distancemedian, maxdist) # let's say under 500 km is close, over 2000km is far
neardist <- geogdist[2] # these two are the output of geogdist above
fardist <- geogdist[3]
chartdist <- c(0, 800, 1200) # this just gets us equal intervals on the graph for the different segments of the lines

plot(geogdist, chartdist)

a <- function(x){ # here is some example code for the piecewise funciton
  ifelse(( x >= 0 & x < neardist),(x * 400/neardist),ifelse((neardist <= x & x < fardist),((x * 400)/(fardist - neardist) + (400-((400)/(fardist - neardist)*neardist))), ifelse((fardist <= x & x <= maxdist),((x * 400)/(maxdist - fardist) + (800-((400)/(maxdist - fardist)*fardist))), NA))) 
} 
plot(a,xlim=c(0,maxdist), ylim = c(0, 1200), col = "red") 

df2$lagrangedistcircstep <- a(df2$distance) # set new lagrange great circle distances using new stepwise function

# function to make new circles with any stepwise function set above
lagrange_predictstep <- function(circlesdataframe) { 
  lagrangecirclesdataframe <- circlesdataframe
  for (row in 1:nrow(circlesdataframe)){
    lagrangecirclesdataframe[row,"r"]<- a(circlesdataframe[row,"r"])}
  return(lagrangecirclesdataframe)
}
lagrangecircles <- lagrange_predictstep(circles) # this projects the circles

```

## 10: Plot with Custom Control Points

```{r}

# Replot coordinates on custom distance scale using piecewise function
df2 <- df2 %>% mutate(
  customdistancex =  (useful::pol2cart(a(distance),ctrPtMathbearing,degrees = TRUE)[[1]]), 
  customdistancey = (useful::pol2cart(a(distance),ctrPtMathbearing,degrees = TRUE)[[2]])
  )

# Combine x and y into a matrix, add as a column, remove x and y columns
# NB: THIS CURRENTLY DOESN"T WORK ***
df2$customcoords <- cbind(df2$customdistancex,df2$customdistancey)
df2 <- select(df2,-starts_with("customdistance"))

# NB: WE SHOULD REALLY SEE ABOUT MAKING A SEPARATE CODE SNIPPET THAT HANDLES ALL OF THE AESTHETICS AT ONCE AND PLUG IN EVERYWHERE

# NB: NEED TO STREAMLINE PLOT NAMES TOO

plot10A <- ggplot(df2, aes( # plot with maxdist/10 circles
                customcoords[,1], 
                customcoords[,2], 
                color = df2$distance)) +  
          geom_circle(aes(x0 = x0, y0 = y0, r = r), 
                      colour = "orange", data = lagrangecircles, 
                      show.legend = NA, inherit.aes = FALSE) + 
          geom_point(stroke = 1, size = df2$valTrans) + 
          geom_point(data = (as.data.frame(ctrPt)), aes(0, 0), color = "orange") + 
          coord_fixed() + 
          labs(color = paste0("Distance from ",ctrPtName," (km)"), x = NULL, y = NULL) + 
          theme(panel.background = element_blank(), 
                axis.ticks = element_blank(), 
                axis.text.x = element_blank(),
                axis.text.y = element_blank())
plot10A

plot10B <- ggplot(df2, aes( # plot without circles, zoomed in on data (do we need x and y lims again to do this programmatically?)
                customcoords[,1], 
                customcoords[,2], 
                color = df2$distance)) + 
          geom_point(stroke = 1, size = df2$valTrans) + 
          geom_point(data = (as.data.frame(ctrPt)), aes(0, 0), color = "orange") + 
          coord_fixed() + 
          labs(color = paste0("Distance from ",ctrPtName," (km)"), x = NULL, y = NULL) + 
          theme(panel.background = element_blank(), 
                axis.ticks = element_blank(), 
                axis.text.x = element_blank(),
                axis.text.y = element_blank())
plot10B

```

# 11a contour test!

```{r}
#LAWN test (wrapper for turfjs, which we should loot at some day)
# points from scratch
Lcoords <- cbind(X = df2$lagrangedistcircX, Y = df2$lagrangedistcircY)
sp <- SpatialPoints(Lcoords)

lawnPts <- c(df2$lagrangedistcircXdf2$lagrangedistcircY)
lawn_tin()



# HUBRIS
getContourLines(df2$lagrangedistcircX, df2$lagrangedistcircY, df2$valTrans, nlevels=5, criticalRatio = 5)
contoureR::contourLinesR(x = df2$lagrangedistcircX, y = df2$lagrangedistcircY, z = df2$valTrans, nlevels = 10)
# WHAT DOES THIS ERROR MEAN!?
  # Error: Values in 'dm' would result in out of bounds errors in xyz.

getContourLines(df2$lon, df2$lat, df2$val, nlevels=5, criticalRatio = 5)

deldir(x, y, dpl=NULL, rw=NULL, eps=1e-09, sort=TRUE, plotit=FALSE,
       digits=6, z=NULL, zdum=NULL, suppressMsge=FALSE, ...)

deldirX <- df2$lagrangedistcircX
deldirY <- df2$lagrangedistcircY
dxy1 <- deldir(deldirX,deldirY)
dxy3 <- deldir(deldirX,deldirY, dpl = list(ndx=2, ndy=2), 
               rw = c(min(deldirX),
                      max(deldirX),
                      min(deldirY),
                      max(deldirY)))
deldir()
 # rw = c(min(df2$lagrangedistcircX), 
                   #        max(df2$lagrangedistcircX), 
                   #        min(df2$lagrangedistcircY), 
                   #        max(df2$lagrangedistcircY))#,
                   #z = df2$valTrans
```


# 11 Interpolation & Integrated plotting

```{r}

df3 <- NULL
df2Sub <- df2[!(df2$ctrBin == TRUE), ]
  #^ for removing nulls
df3 <- with(df2Sub, akima::interp(x = df2Sub$customcoords[,1], y = df2Sub$customcoords[,2], z = val))


# Base R version
plot11A <- filled.contour(x = df3$x,
               y = df3$y,
               z = df3$z,
               color.palette = colorRampPalette(c("white", "blue")),
               main = "value") #+
             #  xlim(-2000,2000) + 
              # ylim(-2000,2000)
plot11A

#########

# ggplot version

df3_grid_df <- melt(df3$z, na.rm = TRUE)
names(df3_grid_df) <- c("x", "y", "val")
df3_grid_df$ldcX <- df3$x[df3_grid_df$x]
df3_grid_df$ldcY <- df3$y[df3_grid_df$y]

plot11B <- ggplot(data = df3_grid_df, 
                 aes(x = ldcX, 
                     y = ldcY, 
                     z = val)) +
          geom_tile(aes(fill = val)) +
          stat_contour(bins = 15, 
                       color = "black", 
                       size = 0.1) +
          scale_fill_gradient(name = "val", 
                              low = "yellow", 
                              high = "red")

plot11B

############################################  
############################################

# Integrated plots, incl sec 10

plot11C <- ggplot() +
          geom_circle(aes(x0 = x0, y0 = y0, r = r), 
                      colour = "#eeeeee", 
                      data = lagrangecircles,
                      size = 0.5,
                      show.legend = NA, 
                      inherit.aes = FALSE) +        
          geom_contour(data = df3_grid_df, 
                 aes(x = ldcX, 
                     y = ldcY, 
                     z = df3_grid_df$val,
                     color = stat(level)),
                 bins = 6,  
                 size = 0.5) +
          scale_colour_gradient(low = "cornflowerblue", high = "red") +
  ### begin simulated drop shadow
            geom_text(data = df2Sub,
                    aes(customcoords[,1],
                         customcoords[,2],
                        label= df2Sub$valName),
                    size = 1.8,
                    check_overlap = TRUE,
                    color = "white") +
            geom_text(data = df2Sub,
                    aes(customcoords[,1],
                         customcoords[,2],
                        label= df2Sub$valName),
                    size = 2,
                    check_overlap = TRUE,
                    color = "white") +
  ### end simulated drop shadow
            geom_text(data = df2Sub,
                    aes(customcoords[,1],
                         customcoords[,2],
                        label= df2Sub$valName,
                        color = df2Sub$valTrans),
                    size = 1.5,
                    check_overlap = TRUE) +
            geom_point(data = (as.data.frame(ctrPt)), 
                     aes(ctrPt[1], ctrPt[2]), 
                     color = "orange", 
                     shape = 8) +
             geom_text(data = (as.data.frame(ctrPt)), 
                   geom = "text", 
                   x = ctrPt[1], y = ctrPt[2], 
                   label = df2$valName[df2$ctrBin == TRUE],
                   color = "orange",
                   size = 2,
                   vjust = 1.25,
                   hjust = -0.25) +
          theme(panel.background = element_blank(), 
                axis.ticks = element_blank(), 
                axis.text.x = element_blank(),
                axis.text.y = element_blank(),
                axis.title = element_blank(),
                legend.position = 'none')
         # scale_fill_gradient(name = "val", low = "yellow", high = "red")

plot11C


plot11D <- ggplot() +
          geom_circle(aes(x0 = x0, y0 = y0, r = r), 
                      colour = "#eeeeee", 
                      data = lagrangecircles,
                      size = 0.5,
                      show.legend = NA, 
                      inherit.aes = FALSE) +        
          geom_contour(data = df3_grid_df, 
                 aes(x = ldcX, 
                     y = ldcY, 
                     z = df3_grid_df$val,
                     color = stat(level)),
                 bins = 6,  
                 size = 0.5) +
          scale_colour_gradient(low = "cornflowerblue", high = "red") +
          geom_point(data = df2Sub,
                aes(customcoords[,1],
                    customcoords[,2],
                    label= df2Sub$valName,
                    alpha = -(df2Sub$lagrangedistcirc)),
                size = .3,
                color="gray35") +
          geom_point(data = (as.data.frame(ctrPt)), 
                     aes(ctrPt[1], ctrPt[2]), 
                     color = "orange", 
                     shape = 8) +
         geom_text(data = (as.data.frame(ctrPt)), 
                   geom = "text", 
                   x = ctrPt[1], y = ctrPt[2], 
                   label = df2$valName[df2$ctrBin == TRUE],
                   color = "orange",
                   size = 2,
                   vjust = 1.25,
                   hjust = -0.25) +
          theme(panel.background = element_blank(), 
                axis.ticks = element_blank(), 
                axis.text.x = element_blank(),
                axis.text.y = element_blank(),
                axis.title = element_blank(),
                legend.position = 'none')
         # scale_fill_gradient(name = "val", low = "yellow", high = "red")

plot11D

```