---
title: "PCP_alphaWorkflow"
author: "VanMcG"
date: "January 26, 2019"
output: html_document
---

## 00: R Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sp)
library(ggplot2)
library(ggforce) # to plot circles
library(dplyr)
#library(devtools)
#library(SpatialEpi)
library(gmt) # geodist is in geosphere
library(tidyverse)
#library(sf)
library(geosphere)
library(useful)
library(MonoPoly)
library(akima)
library(polynom)
library(xkcd) # I don't know if it was my setup or a differnet problem, but had to reinstall ggplot and rlang a few times, so be wary--also, easy enough to just download the font separately
library(extrafont)
library(reshape2)

# USE TO DOWNLOAD XKCD FONT
download.file("http://simonsoftware.se/other/xkcd.ttf", dest="xkcd.ttf", mode="wb")
system("mkdir ~/.fonts")
system("cp xkcd.ttf  ~/.fonts")
font_import(pattern = "xkcd", prompt=FALSE)
# fonts()
# fonttable()

```

## 01: Input CSV into R
* Read and give headers and datatypes for each (eventually do more sophisticated validation/type detection)

```{r}
routes <- read.csv(file = "testData/routes_bare_sm.csv")

routes <- routes[!(routes$lat>50 | routes$long < -125),]
summary(routes)
class(routes) # test to make sure it's a data frame

# at some point, we need to enforce generic column headers

typeof(routes$isDTW)

     # obnoxiously explicit validation

if (typeof(routes$airport) != "character")
  routes$airport <- as.character(routes$airport)
if (typeof(routes$isDTW) != "logical")
  routes$isDTW <- as.logical(routes$isDTW)
if (typeof(routes$lat) != "double")
  routes$isDTW <- as.double(routes$lat)
if (typeof(routes$long) != "double")
  routes$isDTW <- as.double(routes$long)
if (typeof(routes$flightsDTW) != "integer" && typeof(routes$flightsDTW) != "double" )
  routes$flightsDTW <- as.double(routes$flightsDTW)

typeof(routes$isDTW)

```

## 02: Plot test

```{r}
# coordinate validation
ifelse(max(routes$lat > 90.0) || min(routes$lat < -90.0) || max(routes$long > 180.0) || min(routes$long < -180.0), print("Check your coordinate system..."), print("Coordinates look ok!"))

# test plot
plot(routes$long, routes$lat)
```

## 03: Select origin point from rows

* column "iscenter" value = 1
    + __currently pre-set in test data__
* validate that there is only one center

```{r}
if (sum(routes$isDTW) > 1)
  print("Too many centers. Check data")
if (sum(routes$isDTW) == 0)
  print("No center. Check data. Maybe we'll have an interface to select one later")
if (sum(routes$isDTW) == 1)
  print("There is one center! Hooray!")

ctrX <- routes$long[routes$isDTW == TRUE]
ctrY <- routes$lat[routes$isDTW == TRUE]
ctrPt <- c(ctrX,ctrY)
cat("Your center point has the following coordinates:", ctrPt[1],"longitude and", ctrPt[2], "latitude. ")

```

## 04: Calculate all great circle distances and bearings to center point

* __NB:__ half of circumference of earth is ~20,000 km, which might be half of the max dist we should allow
    + example data max is 7,241 km

```{r}
# create a column in routes, fill with km distances
  routes$distance <- geodist(ctrY, ctrX, routes$lat, routes$long, units = "km")  
  maxdist <- max(routes$distance) # max great circle distance

  print( # print out both the longest axial and great circle distances
    paste0("The longest great circle distance between your center point, ",
           routes$airport[routes$isDTW == TRUE], ", and farthest point, ",
           routes$airport[routes$distance == max(routes$distance)], ", is ",
           round(max(routes$distance), 0), " km.")
    )
  
for (row in 1:nrow(routes)){ # convert to polar coordinates
  airportlong <- routes[row,"long"]
  airportlat <- routes[row,"lat"]
  airportlonglat <- c(airportlong, airportlat)
  routes[row,"ctrPtGeobearing"] <- bearing(ctrPt, airportlonglat, a=6378137, f=1/298.257223563)
}
  rm(airportlat,airportlong,airportlonglat,row)
  
for (row in 1:nrow(routes)) # convert from geographic bearings to polar coordinates
  {if(routes$ctrPtGeobearing[row] <= 0)
  # if geobearing is 0 or negative, mathbearing is 90 plus bearing
  routes$ctrPtMathbearing[row] <- abs(routes$ctrPtGeobearing[row]) + 90
  if(routes$ctrPtGeobearing[row] > 0 & routes$ctrPtGeobearing[row] < 90)
  # if geobearing is positive and equal to or under 90, mathbearing is 90 - bearing
  routes$ctrPtMathbearing[row] <- (90 - routes$ctrPtGeobearing[row])
  if(routes$ctrPtGeobearing[row] >= 90 & routes$ctrPtGeobearing[row] <= 180)
  # if geobearing is positive and between 90 and 180 (inclusive), mathbearing is 360 - (geobearing - 90)
  routes$ctrPtMathbearing[row] <- 360 - (routes$ctrPtGeobearing[row] - 90)}
  rm(row)

  routes <- select(routes,-starts_with("ctrPtGeo"))
  # later we use the ctrPtMathbearing and the transformed distances to get back to cartesian coordinates

```

## 05: Initial plot
* GGplot of the lat and long with low-res world background

```{r}
# first plot with ctrPt and data visualized, still using lat/long
plot05 <- ggplot(routes, 
       aes(routes$long, 
           routes$lat, 
           color = routes$distance)) + 
           geom_point() + 
           geom_point(data = (as.data.frame(ctrPt)), aes(ctrPt[1], ctrPt[2]), color = "orange")

plot05

```  
  
## 06: Plot 2 - DTW connectivity
* Plot witn DTW connectivity
  
```{r} 
routes$DTWconnectivity <- (routes$flightsDTW * 1) + .5 

for (row in 1:nrow(routes)){
  routes[row,"circdistancex"] <- (pol2cart(routes$distance[row], routes$ctrPtMathbearing[row], degrees = TRUE))[1]
  routes[row,"circdistancey"] <- (pol2cart(routes$distance[row], routes$ctrPtMathbearing[row], degrees = TRUE))[2]
}
rm(row)

# Lets make some circles
circles <- data.frame(
  x0 = 0,
  y0 = 0,
  r = seq(500, 3500, length.out = 7)
)

# Use coord_fixed to ensure true circularity
ggplot() +
  geom_circle(aes(x0 = x0, y0 = y0, r = r), colour = "orange", data = circles, show.legend = NA, inherit.aes = FALSE) +
  coord_fixed()

plot06 <- ggplot(routes, aes(
                routes$circdistancex, 
                routes$circdistancey, 
                color = routes$distance)) +
        geom_circle(aes(x0 = x0, y0 = y0, r = r),
          colour = "orange", 
          data = circles, 
          show.legend = NA, 
          inherit.aes = FALSE) +
        geom_point(stroke = 1, size = routes$DTWconnectivity) +
        geom_point(data = (as.data.frame(ctrPt)), aes(0, 0), color = "orange") +
        coord_fixed() + labs(color = "Distance in km", x = NULL, y = NULL) +
        guides(colour = "colorbar",size = "legend") +
        theme(panel.background = element_blank())

plot06

```  

## 07: Plot 3 - Log scale
* Plot witn log scale
  
```{r} 
# replot on log scale

routes$distancelog <- log(routes$distance)

for (row in 1:nrow(routes)){
  routes[row,"distancelogcircx"] <- (pol2cart(routes$distancelog[row], routes$ctrPtMathbearing[row], degrees = TRUE))[1]
  routes[row,"distancelogcircy"] <- (pol2cart(routes$distancelog[row], routes$ctrPtMathbearing[row], degrees = TRUE))[2]
}

logcircles <- data.frame()

for (row in 1:nrow(circles)){
  logcircles[row,"r"] <- log(circles[row,"r"])
  logcircles[row,"x0"] <- circles[row,"x0"]
  logcircles[row,"y0"] <- circles[row,"y0"]
}
rm(row)

plot07 <- ggplot(routes, aes(
                routes$distancelogcircx, 
                routes$distancelogcircy, 
                color = routes$distance)) + 
  geom_circle(aes(x0 = x0, y0 = y0, r = r), 
                      colour = "orange", data = logcircles, 
                      show.legend = NA, inherit.aes = FALSE) +
  geom_point(stroke = 1, size = routes$DTWconnectivity) + 
  geom_point(data = (as.data.frame(ctrPt)), aes(0, 0), color = "orange") +
  coord_fixed() + labs(color = "Great circle distance in km", x = NULL, y = NULL) +
  theme(panel.background = element_blank(), axis.ticks = element_blank(),
                axis.text.x = element_blank(),axis.text.y = element_blank())

plot07

```  

## 08 Plot 4 - Square root scale
* Plot witn square root scale
  
```{r} 
routes$distancesqrt <- sqrt(routes$distance)

for (row in 1:nrow(routes)){
  routes[row,"sqrtcircX"] <- (pol2cart(routes$distancesqrt[row], routes$ctrPtMathbearing[row], degrees = TRUE))[1]
  routes[row,"sqrtcircY"] <- (pol2cart(routes$distancesqrt[row], routes$ctrPtMathbearing[row], degrees = TRUE))[2]
}
rm(row)

plot08 <- ggplot(routes, aes(
                routes$sqrtcircX, 
                routes$sqrtcircY, 
                color = routes$distance)) +  
          geom_circle(aes(x0 = x0, y0 = y0, r = r), 
                      colour = "orange", data = sqrt(circles), 
                      show.legend = NA, inherit.aes = FALSE) + 
          geom_point(stroke = 1, size = routes$DTWconnectivity) + 
          geom_point(data = (as.data.frame(ctrPt)), aes(0, 0), color = "orange") + 
          coord_fixed() + 
          labs(color = "Distance in km", x = NULL, y = NULL) + 
          theme(panel.background = element_blank(), axis.ticks = element_blank(),
                axis.text.x = element_blank(),axis.text.y = element_blank())

plot08

```

## 09: Manual Control Points
* Define a decay function using manual control points

```{r}
# polynom::poly
# code to find a Lagrange interpolation polynomial for our control points. for now I'm just defining some, later we should explore the user input and/or data-driven methods.

geogdist <- c(0, 500, 2000, maxdist) # let's say under 500 km is close, over 2000km is far
chartdist <- c(0, 400, 800, 1200) # this just gets us equal intervals on the graph for the different segments of the lines

dat <- data.frame(cbind(geogdist, chartdist)) # chart points to check intervals
ggplot(dat, aes(x=geogdist, y=chartdist)) + geom_point(size=5, col='blue') 

lagrangepoly <- poly.calc(geogdist, chartdist) # calculate polynomial from control points
lagrangepoly # print polynomial 

curve(predict(lagrangepoly,x), from = 0, to = maxdist, n = 100)

predict(lagrangepoly, 200) # print predicted value for geogdist of 200

routes$lagrangedistcirc <- predict(lagrangepoly, routes$distance) # predict great circle distances

# function to make new circles with any lagrange polynomial set above
lagrange_predict <- function(circlesdataframe) { 
  lagrangecirclesdataframe <- circlesdataframe
  for (row in 1:nrow(circlesdataframe)){
    lagrangecirclesdataframe[row,"r"]<- predict(lagrangepoly, circlesdataframe[row,"r"])}
  return(lagrangecirclesdataframe)
}

lagrangecircles <- lagrange_predict(circles) # check if it worked
lagrangecircles

```

## 11A: Plot 1 with manual control points (square)

```{r}

routes$lagrangedistcircX <- NA
routes$lagrangedistcircY <- NA

for (row in 1:nrow(routes)){
  routes[row,"lagrangedistcircX"] <- (pol2cart(routes$lagrangedistcirc[row], routes$ctrPtMathbearing[row], degrees = TRUE))[1] # returns x coordinate only
  routes[row,"lagrangedistcircY"] <- (pol2cart(routes$lagrangedistcirc[row], routes$ctrPtMathbearing[row], degrees = TRUE))[2] # returns y coordinate only
}

plot6a <- ggplot(routes, aes(
                lagrangedistcircX, 
                lagrangedistcircY, 
                color = routes$distance)) + 
          geom_point(stroke = 1, size = routes$DTWconnectivity) + 
          geom_point(data = (as.data.frame(ctrPt)), aes(0, 0), color = "orange") + 
          coord_fixed() + 
          labs(color = "Great circle distance in km") + 
          xlim(-1200,1200) + 
          ylim(-1200,1200)

plot6a

plot6b <- ggplot(routes, aes(
                lagrangedistcircX, 
                lagrangedistcircY, 
                color = routes$distance)) +  
          geom_circle(aes(x0 = x0, y0 = y0, r = r), 
                      colour = "orange", data = lagrange_predict(circles), 
                      show.legend = NA, inherit.aes = FALSE) + 
          geom_point(stroke = 1, size = routes$DTWconnectivity) + 
          geom_point(data = (as.data.frame(ctrPt)), aes(0, 0), color = "orange") + 
          coord_fixed() + 
          labs(color = "Distance in km", x = NULL, y = NULL) + 
          theme(panel.background = element_blank(), 
                axis.ticks = element_blank(), 
                axis.text.x = element_blank(),
                axis.text.y = element_blank())

plot6b


plot7 <- ggplot(routes, aes(
                lagrangedistcircX, 
                lagrangedistcircY, 
                color = routes$distance)) + 
          geom_point(stroke = 1, size = routes$DTWconnectivity) + 
          geom_point(data = (as.data.frame(ctrPt)), aes(0, 0), color = "orange") + 
          coord_fixed() + 
          labs(color = "Great circle distance in km") + 
          xlim(-1200,700) + 
          ylim(-850,600)

plot7

```

## 11A: Plot 1 with manual control points (square)

```{r}
plot6 <- ggplot(routes, aes(
                routes$lagrangedistX * longDiffSign, 
                routes$lagrangedistY * latDiffSign, color = routes$distance)) +
        geom_point(stroke = 1, size = routes$DTWconnectivity) + 
        geom_point(data = (as.data.frame(ctrPt)), aes(0, 0), color = "orange") + 
        coord_fixed() + 
        labs(color = "Great circle distance in km") + 
        xlim(-1200,1200) + 
        ylim(-1200,1200)

plot6
```

## 11B: Plot 2 with manual control points (cropped)

```{r}

plot7 <- ggplot(routes, aes(
                routes$lagrangedistX * longDiffSign, 
                routes$lagrangedistY * latDiffSign, 
                color = routes$distance)) + 
          geom_point(stroke = 1, size = routes$DTWconnectivity) + 
          geom_point(data = (as.data.frame(ctrPt)), aes(0, 0), color = "orange") + 
          coord_fixed() + 
          labs(color = "Great circle distance in km") + 
          xlim(-1200,700) + 
          ylim(-850,600)

plot7
```

## 12: Making maps, Breaking shit

```{r}
### STYLE TEST from plot7


aesX <- routes$lagrangedistX * routes$longDiffSign
aesY <- routes$lagrangedistY * routes$latDiffSign

plot7a <- ggplot(routes, aes(aes(routes$lagrangedistX * longDiffSign, routes$lagrangedistY * latDiffSign, color = routes$distance))) +
             # geom_tile(aes(fill = routes$flightsDTW)) +
              #  + geom_point(stroke = 1, size = routes$DTWconnectivity)  %>%
              geom_text(routes$airport) +
               #stat_contour() +
               #geom_contour(binwidth = 1) +
               labs(x="long", y="lat") +
               xlim(-1200,700) + ylim(-850,600) +
            #scale_fill_continuous(name = "connectivity", low = "white", high = "blue" )
plot7a

# aes(routes$lagrangedistX * longDiffSign, 
#                    routes$lagrangedistY * latDiffSign, 
#                    color = routes$distance)


plot7b <- ggplot(routes, aes(aesX, aesY, 
                   color = routes$distance)) + 
  geom_text(aes(label=routes$airport), size = 2, check_overlap = TRUE) + 
  geom_point(data = (as.data.frame(ctrPt)), aes(0, 0), color = "orange") + 
  coord_fixed() + 
  labs(color = "Great circle distance in km") + 
  xlim(-1200,700) + 
  ylim(-850,600) +
  theme_bw() +
  theme(
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.text.y = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

plot7b


plot1a <- ggplot(routes, 
                aes(routes$long, routes$lat, 
                    color = routes$distance)) + 
  geom_text(aes(label=routes$airport), size = 2, family= "xkcd", check_overlap = TRUE) + 
  geom_point(data = (as.data.frame(ctrPt)), 
             aes(ctrPt[1], ctrPt[2]), color = "orange") +
  xlim(xlim1,xlim2) +
  ylim(ylim1,ylim2) +
  labs(x="long", y="lat") +
  labs(color = "distance") +
  theme_xkcd() +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )


plot1a
```

interpolation experiments
```{r}

flts <- NULL
flts <- with(routes, interp(x = lagrangedistcircX, y = lagrangedistcircY, z = flightsDTW))

# Base R version
contourEX <- filled.contour(x = flts$x,
               y = flts$y,
               z = flts$z,
               color.palette = colorRampPalette(c("white", "blue")),
               main = "DTW Flights") +
               xlim(-2000,2000) + 
               ylim(-2000,2000)
contourEX

#########
plot8 <- ggplot(flts)


#################
f_df <- data.frame(f_z = flts$z,
                   f_x = flts$x,
                   f_y = flts$y)
        # f_df is very strange looking. Not sure what's going on there

f_df <- data.frame(flts$z, flts$x, flts$y)

ggplot(data = f_df, aes(x = f_df$x, y = f_df$y, z = f_df$z)) +
  stat_contour()

f_df <- melt(f_df, id = "f_z")

# GGPLOT2
f_df <- melt(flts$z, na.rm = TRUE)
names(df) <- c("x", "y", "z")
f_df$X <- flts$x[flts$x]
f_df$Y <- flts$y[flts$y]

ggplot(data = f_df, aes(x = X, y = Y, z = Z)) +
  geom_tile(aes(fill = flights)) +
  stat_contour()

```

# TEST FOR NEW LAGRANGE METHODS

```{r} 

dat <- data.frame(cbind(geogdist, chartdist)) # chart points to check intervals
lagrangemonpol <- monpol(formula = chartdist~geogdist, data = dat, plot.it = TRUE,
                         ptype = "SOS", monotone = "increasing")

lagrangemonpol
predict(lagrangemonpol,100)

predict(lagrangemonpol, c(0,100,200,300,400,500,600), scale = c("original"))

curve(predict(lagrangemonpol,x), from = 0, to = maxdist, n = 4)

print(lagrangemonpol, digits = max(3, getOption("digits") - 3))

# plot(lagrangemonpol) #EM 3/29: this breaks for me. "'x' is a list, but does not have components 'x' and 'y'"

median(routes[["distance"]])
geogdist <- c(0, 1220.595, maxdist) # let's say under 500 km is close, over 2000km is far
chartdist <- c(0, 600, 1200) # this just gets us equal intervals on the graph for the different segments of the lines

# abline(lm(chartdist~geogdist)) #EM 3/29: breaks for me, possibly due to line 450

```
